Effect of multiple reflection of the sky signal off the HERA element on the measurements of 21cm power spectrum

### 2015-11-02 DishReflectometry/Paper1_v0.tex
Monday update

\subsection{Visibility of a two element interferometer considering reflections to first order}
Upon first incidence of the sky signal $V_{sky}$, $\Gamma_{f}$ fraction of the voltage is reflected off the feed while $(1-\Gamma_{f})$ fraction is coupled to the feed. The reflected voltage then further reflected off the dish and $(1-\Gamma_{f}) $ fraction of it reenters the feed with a phase delay $\Phi$. Hence, the net voltage at the output of the two antennas could be written as,    
\begin{eqnarray}
V_{1} & = & A V_{sky}(1-\Gamma_{f}) + A V_{sky} \Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau}  (1-\Gamma_{f})^{2}\nonumber\\
         & = & A V_{sky} (1-\Gamma_{f}) \left [ 1+ (1-\Gamma_{a})\Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau} \right]\nonumber\\
\label{eq4}
\end{eqnarray}
\begin{eqnarray}
V_{2} & = & A V_{sky}(1-\Gamma_{f}) + A V_{sky} \Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau} (1-\Gamma_{f})^{2}exp({2\pi i \nu \vec{b}.\hat{s} \over c} )\nonumber\\
         & = & A V_{sky} (1-\Gamma_{f}) \left [ 1+(1-\Gamma_{a}) \Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau}  \right]exp({2\pi i \nu \vec{b}.\hat{s} \over c} )\nonumber\\
\label{eq5}
\end{eqnarray}
Frequency as well as angular dependence of the quantities are not written explicitly for notational simplicity. 
The two antennas are assumed to have identical beam pattern which is $A_{1}=A_{2} =A$. We also absorb the common gain term $(1-\Gamma_{f})$ with antenna response $A$ and denote them as $A_{f}= A(1-\Gamma_{f})$ here after. Hence, the complex visibility could be,

@@

Therefore, the total visibility at any frequency would be the sum of the three visibilities contributed by three sources of correlated power. First term in the square bracket represents the cross coorrelation between the two antenna outputs contributed by the sky signal upon first direct incidence. The second term results from the cross correlation between two reflected signals in each antenna.The last term represents the cross correlation between the first direct incident signal of one antenna and the first reflected signal from the other antenna and vice versa.
If the dish vertex is so constructed such that the reflection from the dish vertex is negligible i.e, $\Gamma_{d}<<1\ or = 0$ then the second and the third term of the $\Sigma$ vanish. \\

@@

\subsection{Effect of first reflection on the delay spectrum}
We define $B(\vec \theta, \nu) =  A^{2}(\vec \theta, \nu) $ as the beam power pattern of the interferometer and $\Sigma_{rr}(\nu), \Sigma_{dr}(\nu)$ are the correlation coefficient between the two reflected signals in each antenna and the direct signal of one antenna with the reflected signal from the other antenna in the uv plane i.e,

\begin{equation}
\Sigma_{rr}(\nu) =  \left [ |(1-\Gamma_{f})|^{2}|\Gamma_{f}|^{2}|\Gamma_{d}|^{2} \right ]
\end{equation}

\begin{equation}
\Sigma_{rr}(\nu) = \left [ 2 \Re \left ((1-\Gamma_{f})\Gamma_{f}\Gamma_{d} e^{i\Phi} \right) \right]
\end{equation}

and $\Delta \tau$ is the additional delay each reflected component suffers in addition to the intrinsic geometric delay 
$\tau_{g}$. Therefore, $\Sigma(\vec \theta, \nu) = B(\vec \theta, \nu) \left [ 1+\Sigma_{rr} +\Sigma_{dr} \right]$

As given in equation \ref{eq3}, the Fourier transform of this visibility spectrum along the frequency axis results in the delay spectrum. Hence, the all sky integrated delay spectrum could be written as, 
%\begin{eqnarray}
%\mathcal{F} \left [V e^{i\phi}\right] & = & \int \left [\int B I_{sky}exp({2\pi i \nu \vec{b}.\hat{s} \over c} ) d\Omega \right ] exp\left[ -2 \pi i\nu \tau \right] d\nu \nonumber\\
%&& + \int \left [  \int B I_{sky} \Sigma_{rr} exp({2\pi i \nu \vec{b}.\hat{s} \over c} ) d\Omega \right] exp\left[ -2 \pi i\nu \tau \right] d\nu  \nonumber\\
%&& +\int \left [ \int B I_{sky} \Sigma_{dr} exp({2\pi i \nu \vec{b}.\hat{s} \over c} ) d\Omega \right ] exp\left[ -2 \pi i\nu \tau \right] d\nu \nonumber\\	
%&& + \int \left [  \int |A_{f}|^{2}|V_{sky}|^{2} |(1-\Gamma_{f})|^{2}|\Gamma_{f}|^{2}|\Gamma_{d}|^{2} exp({2\pi i \nu \vec{b}.\hat{s} \over c} ) d\Omega \right] exp\left[ -2 \pi i\nu \tau \right] d\nu  \nonumber\\
%&& +\int \left [ \int |A_{f}|^{2}|V_{sky}|^{2} 2 \Re \left ((1-\Gamma_{f})\Gamma_{f}\Gamma_{d} e^{i\Phi} \right)exp({2\pi i \nu \vec{b}.\hat{s} \over c} ) d\Omega \right ] exp\left[ -2 \pi i\nu \tau \right] d\nu \nonumber\\			
%			& = & \int \left [\int B I_{sky} exp({2\pi i \nu \tau_{g}} ) dl dm \right ] exp\left[ -2 \pi i\nu \tau \right] d\nu \nonumber\\
%&& + \int \left [  \int B I_{sky} \Sigma_{rr} exp({2\pi i \nu \tau_{g}} ) dl dm \right] exp\left[ -2 \pi i\nu \tau \right] d\nu  \nonumber\\
%&& +\int \left [ \int B I_{sky} \Sigma_{dr} exp({2\pi i \nu \tau_{g}} ) dl dm \right ] exp\left[ -2 \pi i\nu \tau \right] d\nu \nonumber\\
%\label{eq9}
%\end{eqnarray}
%
\begin{eqnarray}			
\mathcal{F} \left [V e^{i\phi}\right] & = & \left [ B(\tau)\ast I_{sky}(\tau) \right ]\ast \delta(\tau+\tau_{g}) \nonumber\\
		&& +  \left [ B(\tau)\ast I_{sky}(\tau)* \Sigma_{rr}( \tau) \right ]\ast \delta(\tau+\tau_{g}) \nonumber\\
		&& +  \left [B(\tau)\ast I_{sky}(\tau)* \Sigma_{dr}( \tau) \right ]\ast \delta(\tau+\tau_{g}+ \Delta \tau)\nonumber\\
\label{eq10}
\end{eqnarray}

 In words, the Fourier transform of the visibility contributed by any two voltage components from two antennas with no mutual delay is centred at a delay $\tau = \tau_{g}$ in the delay spectrum whereas any two voltage components from two antennas components having a mutual delay of $\Delta \tau$ will be centred at $\tau +\Delta \tau$.

>>>>

\subsection{Effect of first order reflections on visibility and delay spectrum}
Upon first incidence of the sky signal $V_{sky}$, $\Gamma_{f}$ fraction of the voltage is reflected off the feed while $(1-\Gamma_{f})$ fraction is coupled to the feed. The reflected voltage then further reflected off the dish and $(1-\Gamma_{f}) $ fraction of it reenters the feed with a roundtrip time delay $\Delta \tau$. Hence, the net voltage at the output of the two antennas could be written as,    
\begin{eqnarray}
V_{1} & = & A V_{sky}(1-\Gamma_{f}) + A V_{sky} \Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau}  (1-\Gamma_{f})^{2}\nonumber\\
         & = & A V_{sky} (1-\Gamma_{f}) \left [ 1+ (1-\Gamma_{a})\Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau} \right]\nonumber\\
\label{eq4}
\end{eqnarray}
\begin{eqnarray}
V_{2} & = & A V_{sky}(1-\Gamma_{f}) + A V_{sky} \Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau} (1-\Gamma_{f})^{2}exp({2\pi i \nu \vec{b}.\hat{s} \over c} )\nonumber\\
         & = & A V_{sky} (1-\Gamma_{f}) \left [ 1+(1-\Gamma_{a}) \Gamma_{f} \Gamma_{d} e^{2\pi i\nu \Delta \tau}  \right]exp({2\pi i \nu \vec{b}.\hat{s} \over c} )\nonumber\\
\label{eq5}
\end{eqnarray}
Frequency as well as angular dependence of the quantities are not written explicitly for notational simplicity. 
The two antennas are assumed to have identical beam pattern which is $A_{1}=A_{2} =A$. We also absorb the common gain term $(1-\Gamma_{f})$ with antenna response $A$ and denote them as $A_{f}= A(1-\Gamma_{f})$ here after. Hence, the complex visibility could be,

@@

Instrumental response function acquires additional terms arising due to correlation between the directed and reflected component of the sky signal. The total visibility at any frequency would be the sum of the three visibilities contributed by three sources of correlated power. First term in the square bracket represents the cross correlation between the two antenna outputs contributed by the sky signal upon first direct incidence. The second term results from the cross correlation between two reflected signals in each antenna.The last term represents the cross correlation between the first direct incident signal of one antenna and the first reflected signal from the other antenna and vice versa.

Fourier transform of this visibility spectrum along the frequency axis results in the delay spectrum. 
 Fourier transform of the visibility contributed by any two voltage components from two antennas with no mutual delay is centred at a delay $\tau = \tau_{g}$ in the delay spectrum whereas any two voltage components from two antennas components having a mutual delay of $\Delta \tau$ will be centred at $\tau_{g} +\Delta \tau$. Therefore, the correlation between the reflected signal  from the feed and subsequently from the dish with the direct signal thus extends the foreground response to delays beyond the maximum geometric delay for a given baseline and contaminates the delay mode where 21~cm delay spectrum could potentially be detected. Also, any reflections arriving at the feed at any non-zero  value of the delay $\Delta \tau$ which could be the result of reflections from the feed structures would also generate non zero delay response at those delays. 
If the dish vertex is so constructed such that the reflection from the dish vertex is negligible i.e, $\Gamma_{d}<<1\ or = 0$ then the second and the third term of the $\Sigma$ vanish. Also, if we assume a $\Gamma_{a} =-10dB$ and $\Gamma_{d}= -20dB$, the second term in the expression of $\Sigma(\vec \theta, \nu)$ becomes negligible compared to the first and the third term of the expression and, therefore, can be neglected. \\

@@

%\subsection{Effect of first reflection on the delay spectrum}
%We define $B(\vec \theta, \nu) =  A^{2}(\vec \theta, \nu) $ as the cross power pattern of the interferometer and $\Sigma_{rr}(\nu), \Sigma_{dr}(\nu)$ are the correlation coefficient between the two reflected signals in each antenna and the direct signal of one antenna with the reflected signal from the other antenna in the uv plane i.e,

%\begin{equation}
%\Sigma_{rr}(\nu) =  \left [ |(1-\Gamma_{f})|^{2}|\Gamma_{f}|^{2}|\Gamma_{d}|^{2} \right ]
%\end{equation}

%\begin{equation}
%\Sigma_{dr}(\nu) = \left [ 2 \Re \left ((1-\Gamma_{f})\Gamma_{f}\Gamma_{d} e^{i\Phi} \right) \right]
%\end{equation}

%and $\Delta \tau$ is the additional delay each reflected component suffers in addition to the intrinsic geometric delay 
%$\tau_{g}$. Therefore, $\Sigma(\vec \theta, \nu) = B(\vec \theta, \nu) \left [ 1+\Sigma_{rr} +\Sigma_{dr} \right]$

%As given in equation \ref{eq3}, the Fourier transform of this visibility spectrum along the frequency axis results in the delay spectrum. Hence, the all sky integrated delay spectrum could be written as, 
%\begin{eqnarray}
%\mathcal{F} \left [V e^{i\phi}\right] & = & \int \left [\int B I_{sky}exp({2\pi i \nu \vec{b}.\hat{s} \over c} ) d\Omega \right ] exp\left[ -2 \pi i\nu \tau \right] d\nu \nonumber\\
%&& + \int \left [  \int B I_{sky} \Sigma_{rr} exp({2\pi i \nu \vec{b}.\hat{s} \over c} ) d\Omega \right] exp\left[ -2 \pi i\nu \tau \right] d\nu  \nonumber\\
%&& +\int \left [ \int B I_{sky} \Sigma_{dr} exp({2\pi i \nu \vec{b}.\hat{s} \over c} ) d\Omega \right ] exp\left[ -2 \pi i\nu \tau \right] d\nu \nonumber\\	
%&& + \int \left [  \int |A_{f}|^{2}|V_{sky}|^{2} |(1-\Gamma_{f})|^{2}|\Gamma_{f}|^{2}|\Gamma_{d}|^{2} exp({2\pi i \nu \vec{b}.\hat{s} \over c} ) d\Omega \right] exp\left[ -2 \pi i\nu \tau \right] d\nu  \nonumber\\
%&& +\int \left [ \int |A_{f}|^{2}|V_{sky}|^{2} 2 \Re \left ((1-\Gamma_{f})\Gamma_{f}\Gamma_{d} e^{i\Phi} \right)exp({2\pi i \nu \vec{b}.\hat{s} \over c} ) d\Omega \right ] exp\left[ -2 \pi i\nu \tau \right] d\nu \nonumber\\			
%			& = & \int \left [\int B I_{sky} exp({2\pi i \nu \tau_{g}} ) dl dm \right ] exp\left[ -2 \pi i\nu \tau \right] d\nu \nonumber\\
%&& + \int \left [  \int B I_{sky} \Sigma_{rr} exp({2\pi i \nu \tau_{g}} ) dl dm \right] exp\left[ -2 \pi i\nu \tau \right] d\nu  \nonumber\\
%&& +\int \left [ \int B I_{sky} \Sigma_{dr} exp({2\pi i \nu \tau_{g}} ) dl dm \right ] exp\left[ -2 \pi i\nu \tau \right] d\nu \nonumber\\
%\label{eq9}
%\end{eqnarray}
%
%\begin{eqnarray}			
%\mathcal{F} \left [V e^{i\phi}\right] & = & \left [ B(\tau)\ast I_{sky}(\tau) \right ]\ast \delta(\tau+\tau_{g}) \nonumber\\
%		&& +  \left [ B(\tau)\ast I_{sky}(\tau)* \Sigma_{rr}( \tau) \right ]\ast \delta(\tau+\tau_{g}) \nonumber\\
%		&& +  \left [B(\tau)\ast I_{sky}(\tau)* \Sigma_{dr}( \tau) \right ]\ast \delta(\tau+\tau_{g}+ \Delta \tau)\nonumber\\
%\label{eq10}
%\end{eqnarray}

 %In words, the Fourier transform of the visibility contributed by any two voltage components from two antennas with no mutual delay is centred at a delay $\tau = \tau_{g}$ in the delay spectrum whereas any two voltage components from two antennas components having a mutual delay of $\Delta \tau$ will be centred at $\tau_{g} +\Delta \tau$. The correlation between the reflected signal  from the feed and subsequently from the dish with the direct signal thus extends the foreground response to delays beyond the maximum geometric delay for a given baseline and contaminates the delay mode where 21~cm delay spectrum could potentially be detected. Also, any reflections arriving at the feed at any non-zero  value of the delay $\Delta \tau$ which could be the result of reflections from the feed structures would generate non zero delay response at those delays.

\subsection{Effect of higher order reflection on the delay spectrum} 
Correlation between the higher order reflections of the signal in any antenna with the direct component of the other would result in a delay response at integral multiple of $\Delta \tau$ and vice versa. 
If $I_{sky}$ is reflected n times in between the feed and the dish, the net voltage entering the feed after an
$n^{th}$ reflection off the feed and the dish is written as:
\begin{equation}
V_{1} =  A V_{sky}(1-\Gamma_{f})[1+ \Gamma_{f}\Gamma_{d} e^{i\phi}+ (\Gamma_{f}\Gamma_{d})^2e^{i2\phi}+ ....+ (\Gamma_{f}\Gamma_{d})^{n}e^{in\phi}]
\end{equation}
and,

\begin{equation}
V_{2} = A V_{sky}(1-\Gamma_{f})[1+ \Gamma_{f}\Gamma_{d} e^{i\phi}+ (\Gamma_{f}\Gamma_{d})^2e^{i2\phi}+ ....+ (\Gamma_{f}\Gamma_{d})^{n}e^{in\phi}]
\end{equation}

where $\Phi = 2\pi\nu\Delta \tau$. The relevant terms in the corresponding visibility and subsequently in the delay spectrum would be the ones which represent the correlation between the direct component of one antenna with the reflected components of the other resulting in higher order terms  in $\Sigma(\vec \theta, \nu)$. The corresponding instrument response may be written as,

\begin{equation}
\Sigma(\vec \theta, \nu) = A_{f} \left [ 1+ 2 \Re \left [(1-\Gamma_{f})( \Gamma_{f}\Gamma_{d} e^{i\phi}+ (\Gamma_{f}\Gamma_{d})^2e^{i2\phi}+ ....+ (\Gamma_{f}\Gamma_{d})^{n}e^{in\phi}) \right ]\right ]
\end{equation}
